A Deeper Analysis into the Trends and Factors of the US's Declining Birth Rates

### 2024-02-13 paper/paper.qmd
added figure 2f to quarto document

# acquire fig 2 a data from figs_2a_2b.csv
fig_2a_data <-
  read.csv((file=here("data/figs_2a_2b.csv")))

@@

# acquire fig 2 d data from figs_1_2_3.csv
fig_2d_data <-
  read.csv((file=here("data/figs_1_2_3.csv")))

@@

@@

# replicate figure 2d from the original paper
fig_2d_data_long %>% ggplot(aes(x = year, y = Birth_Rate, color = Education_Level)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "Births per 1,000 women in relevant population subgroup", color = "Education Level", caption = "Mother’s level of education (ages 20–44)") +
  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 150, by = 20)) +
  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5)) +
  geom_text(aes(x = 1995, y = 105), label = "No high school degree", color = "black") +
  geom_text(aes(x = 1995, y = 75), label = "College graduate
", color = "#d44215") +
  theme_classic() +
  geom_text(aes(x = 1995, y = 65), label = "High school graduate", color = "#0cc4be") +
  geom_text(aes(x = 1995, y = 45), label = "Some college", color = "black") +
  theme_classic() +
  theme(legend.position = "none")
```

>>>>

# acquire fig 2a data from figs_2a_2b.csv
fig_2a_data <-
  read.csv((file=here("data/figs_2a_2b.csv")))

@@

# acquire fig 2d data from figs_1_2_3.csv
fig_2d_data <-
  read.csv((file=here("data/figs_1_2_3.csv")))

@@

# acquire figure 2F data from NCHS Natality Microdata's nchs_births_pop_1990_2019.dta and SEER's age_race_comp_seer.dta
nchs_births_data <- read_dta(here("data/nchs_births_pop_1990_2019.dta"))
seer_data <- read_dta(here("data/age_race_comp_seer.dta"))

merged_data <- merge(nchs_births_data, seer_data, by = c("stname", "year"))

merged_data <- merged_data %>%
  mutate(
    brate_firstbirth = merged_data$numbirth_firstbirth / merged_data$pop1544 * 1000,
    brate_secondbirth = merged_data$numbirth_secondbirth / merged_data$pop1544 * 1000,
    brate_thirdbirth = merged_data$numbirth_thirdbirth / merged_data$pop1544 * 1000,
    brate_fourthbirth = merged_data$numbirth_fourthbirth / merged_data$pop1544 * 1000
  ) %>%
  select(year, starts_with("brate_"))

merged_data_long <- merged_data %>%
  pivot_longer(cols = -year, names_to = "Parity", values_to = "Birth_Rate")

@@

# replicate figure 2d from the original paper
fig_2d_data_long %>% ggplot(aes(x = year, y = Birth_Rate, color = Education_Level)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "Births per 1,000 women in relevant population subgroup", color = "Education Level", caption = "Mother’s level of education (ages 20–44)") +
  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 150, by = 20)) +
  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5)) +
  geom_text(aes(x = 1995, y = 105), label = "No high school degree", color = "black") +
  geom_text(aes(x = 1995, y = 75), label = "College graduate
", color = "#d44215") +
  theme_classic() +
  geom_text(aes(x = 1995, y = 65), label = "High school graduate", color = "#0cc4be") +
  geom_text(aes(x = 1995, y = 45), label = "Some college", color = "black") +
  theme_classic() +
  theme(legend.position = "none")

# replicate figure 2f from the original paper
merged_data_long %>% ggplot(aes(x = year, y = round(Birth_Rate, digits = 1), color = Parity)) +
  geom_line(size = 1, stat = "summary", fun = mean) +
  labs(x = "Year", y = "Births per 1,000 women in relevant population subgroup", color = "Parity", caption = "Parity (ages 15-44)") +
  scale_y_continuous(limits = c(0, 40), breaks = seq(0, 40, by = 10)) +
  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5)) +
  geom_text(aes(x = 1995, y = 28), label = "First birth", color = "black") +
  geom_text(aes(x = 1995, y = 23), label = "Second birth", color = "black") +
  geom_text(aes(x = 1995, y = 12.5), label = "Third Birth", color = "black") +
  geom_text(aes(x = 1995, y = 8.5), label = "Fourth+ birth", color = "black") +
  theme_classic() +
  theme(legend.position = "none")
```
